Cognate Priming

### 2023-07-10 docs/index.qmd
Poossible coding problem

# load target objects
tar_load(c(participants, stimuli, 
		   attrition_trials,
		   attrition_participants,
		   gaze_aoi,
		   data_time, data_summary))

@@

Trial-level inclusion criteria:

1. Valid prime: $\text{Prime}_{LT} \geq 750 \text{ms}$ (50%)
2. Valid test: $\text{Target}_{LT} + \text{Distractor}_{LT} \geq 1,000 \text{ms}$ (50%)
3. Minimum 100ms looking time to target, and minimum 100ms looking time to target (5%)

Participant-level inclusion criteria:

4. At least two valid trial of each type (cognate, non-cognate, unrelated)

@@

```{r}
#| label: fig-prior-epred
#| fig-width: 9
#| fig-height: 6
nd <- tibble(time_bin = seq(0, 20, 0.1)) |> 
polypoly::poly_add_columns(time_bin, degree = 3, prefix = "ot", scale_width = 1)

m <- add_epred_draws(nd, fit_prior, ndraws = 50, re_formula = NA) |> 
mutate(.epred = plogis(.epred))

ggplot(m, aes(time_bin, .epred)) +
geom_hline(yintercept = 0.5, colour = "grey", size = 1) +
geom_line(aes(group = .draw), colour = clrs[4], size = 1, alpha = 0.5) +
labs(x = "Time (ms)",
y = "Proportion of target looking time") +
scale_y_continuous(labels = percent) +
scale_x_continuous(labels = function(x) format(x*100, big.mark = ","))
```

@@

## Attrition

::: {.tabset-panel}

### By participants

### By stimuli properties

## Descriptive statistics

### Aggregated target looking

```{r}
#| label: fig-descriptives-aggregated
#| fig-width: 10
#| fig-height: 6
data_summary |> 
summarise(.prop = mean(.prop),
.by = c(id, lp, trial_type, age_group)) |> 
ggplot(aes(trial_type, 
.prop, 
colour = trial_type,
fill = trial_type)) +
facet_grid(lp~age_group) +
geom_hline(yintercept = 0.5, 
size = 1, 
colour = "grey") +
geom_line(aes(group = id),
colour = "grey",
alpha = 0.25) +
geom_dots(layout = "swarm") +
geom_boxplot(width = 0.1,
fill = "white", 
colour = "black",
outlier.colour = NA) +
labs(x = "Trial type",
y = "Proportion of target looking",
colour = "Trial type", 
fill = "Trial type") +
scale_y_continuous(labels = \(x) scales::percent(x, accuracy = 1)) +
theme(legend.position = "none")
```

### Time-course target looking

```{r}
#| label: fig-descriptives-time-course
#| fig-width: 12
#| fig-height: 8
data_time |> 
summarise(.logit = mean(.logit),
.by = c(id, timebin, lp, trial_type, age_group)) |> 
ggplot(aes(timebin, .logit,
colour = trial_type, 
fill = trial_type, 
shape = trial_type)) +
facet_grid(lp~age_group) +
geom_hline(yintercept = 0, 
size = 1,
colour = "grey") +
stat_summary(fun.data = mean_se,
geom = "ribbon",
colour = NA,
alpha = 0.5) +
stat_summary(fun = mean,
geom = "point",
size = 2) +
stat_summary(fun = mean,
geom = "line", 
size = 0.75) +
labs(x = "Time bin (100 ms)",
y = "Proportion of target looking",
colour = "Trial type", 
fill = "Trial type", 
shape = "Trial type") +
scale_y_continuous(labels = \(x) scales::percent(x, accuracy = 1)) +
scale_x_continuous(labels = \(x) format(x * 100, big.mark = ",")) +
theme(legend.position = "top",
legend.title = element_blank(),
panel.grid = element_line(linetype = "dotted", 
colour = "grey"),
panel.grid.major = element_line(colour = "grey", 
linetype = "dotted"))
```

## Model outputs

### Model comparison

```{r}
#| label: tab-model-comparison
loos |> 
as_tibble(.name_repair = make_clean_names) |> 
rownames_to_column("model") |> 
relocate(
model, 
matches("diff"),
matches("elpd_loo"),
matches("p_loo"), 
matches("looic")
) |>
mutate(across(everything(), as.numeric),
model = attr(loos, "dimnames")[[1]]) |> 
select(-matches("subsampling")) |> 
gt(rowname_col = "model") |>
fmt_number(is.numeric, decimals = 1) |> 
cols_merge_uncert(col_val = elpd_diff, col_uncert = se_diff) |> 
cols_merge_uncert(col_val = elpd_loo, col_uncert = se_elpd_loo) |> 
cols_merge_uncert(col_val = p_loo, col_uncert = se_p_loo) |> 
cols_merge_uncert(col_val = looic, col_uncert = se_looic) |> 
cols_label(model = "Model",
elpd_loo = md("*ELPD*"),
p_loo = md("*p*"),
looic = md("*LOO-IC*"),
elpd_diff = md("Difference")) |> 
tab_source_note(md("Pareto-*k* estimates of all models were acceptable (*k* < 0.5)")) |> 
tab_style(cell_text(weight = "bold"),
cells_column_labels()) |> 
tab_style(cell_fill(color = "grey", alpha = 0.5),
cells_body(matches("diff"))) |> 
tab_style(cell_fill(color = clrs[4], alpha = 0.75),
cells_body(everything(), 1))
```

### Fixed effects

#### Table

```{r}
#| label: tbl-fixed-effects
#| tbl-cap: "Summary of the posterior distribution of the Growth Curve Analysis model. We report the median and 95% highest density interval (*HDI*) of each fixed coefficient. The 95% interval indicates the range of values that contains the true value of the parameter with 95% probability. Values are in *logit* scale. Coefficients whose 95% HDI excludes zero are highlighted in red. Coefficients on the intercept indicate changes in overal proportion of target looking, coefficients in the linear term indicate changes in the overall slope of target looking growth across time, coefficients in the quadratic and cubic terms reflect the steepness of the rise and fall or of the rise, fall and rise, respectivey,  of the curve."
post_draws <- gather_draws(fit_3, `b_.*`, regex = TRUE)

str_repl <- c(
"b_Intercept" = "Base",
"b_ot1" = "Base",
"b_ot2" = "Base",
"b_ot3" = "Base",
"b_age_group21_25" = "Age [21-25]",
"b_age_group25_30" = "Age [25-30]",
"b_lpmon_bil" = "Group",
"b_trial_typerelated_unrelated" = "Prime (R)",
"b_trial_typecognate_noncognate" = "Prime (C)",

"b_age_group21_25:lpmon_bil" = "Age [21-25] × Group",
"b_age_group25_30:lpmon_bil" = "Age [25-30] × Group",
"b_age_group21_25:trial_typerelated_unrelated" = "Age [21-25] × Prime (R)",
"b_age_group21_25:trial_typecognate_noncognate" = "Age [21-25] × Prime (C)",
"b_age_group25_30:trial_typerelated_unrelated" = "Age [25-30] × Prime (R)",
"b_age_group25_30:trial_typecognate_noncognate" = "Age [25-30] × Prime (C)",
"b_lpmon_bil:trial_typerelated_unrelated" = "Group × Prime (R)",
"b_lpmon_bil:trial_typecognate_noncognate" = "Group × Prime (C)",

"b_age_group21_25:lpmon_bil:trial_typerelated_unrelated" = "Age [21-25] × Group × Prime (R)",
"b_age_group21_25:lpmon_bil:trial_typecognate_noncognate" = "Age [21-25] × Group × Prime (C)",
"b_age_group25_30:lpmon_bil:trial_typerelated_unrelated" = "Age [25-30] × Group × Prime (R)",
"b_age_group25_30:lpmon_bil:trial_typecognate_noncognate" = "Age [25-30] × Group × Prime (C)"
)

post_draws |> 
median_hdi() |>
mutate(poly = case_when(str_detect(.variable, "ot1") ~ "Linear",
str_detect(.variable, "ot2") ~ "Quadratic",
str_detect(.variable, "ot3") ~ "Cubic",
TRUE ~ "Intercept"),
.variable_name = case_when(.variable %in% paste0("b_ot", 1:3) ~ .variable,
TRUE ~ str_remove_all(.variable, "ot1\\:|ot2\\:|ot3\\:")) |> 
factor(levels = names(str_repl),
label = str_repl,
ordered = TRUE),
type = case_when(.variable_name %in% str_repl[1:9] ~ "Main effect",
.variable_name %in% str_repl[10:17] ~ "Two-way interactions",
.variable_name %in% str_repl[18:length(str_repl)] ~ "Three-way interactions")) |>
select(.variable_name, type, poly, .value:.upper) |>
arrange(.variable_name) |> 
pivot_wider(names_from = poly, 
values_from = is.numeric,
id_cols = c(.variable_name, type),
values_fill = NA,
names_repair = make_clean_names) |> 
relocate(variable_name, starts_with("value_")) |> 
gt(rowname_col = "variable_name",
groupname_col = "type") |> 
fmt_number(is.numeric) |>
tab_spanner("Intercept", matches("intercept")) |> 
sub_missing(everything(), rows = everything(), missing_text = "--") |> 
tab_spanner("Linear", matches("linear")) |> 
tab_spanner("Quadratic", matches("quadratic")) |> 
tab_spanner("Cubic", matches("cubic")) |> 
cols_merge(c(lower_intercept, upper_intercept), pattern = "[{1}, {2}]") |> 
cols_merge(c(lower_linear, upper_linear), pattern = "[{1}, {2}]") |> 
cols_merge(c(lower_quadratic, upper_quadratic), pattern = "[{1}, {2}]") |> 
cols_merge(c(lower_cubic, upper_cubic), pattern = "[{1}, {2}]") |>
cols_label(value_intercept = md("b"),
lower_intercept = md("95% *HDI*"),
value_linear = md("b"),
lower_linear = md("95% *HDI*"),
value_quadratic = md("b"),
lower_quadratic = md("95% *HDI*"),
value_cubic = md("b"),
lower_cubic = md("95% *HDI*")) |> 
tab_style(cell_text(align = "left", indent = px(10)),
cells_stub()) |> 
tab_style(cell_fill(color = "grey90"),
cells_body(matches("value"))) |> 
tab_style(cell_text(weight = "bold"),
cells_column_spanners()) |> 
tab_style(cell_text(weight = "bold"),
cells_row_groups()) |> 
tab_style(cell_text(style = "oblique"),
cells_column_labels(columns = matches("value"))) |> 
tab_style(cell_fill(color = clrs[4], alpha = 0.75),
cells_body(columns = value_intercept, 
rows = (upper_intercept < 0) | (lower_intercept > 0))) |> 
tab_style(cell_fill(color = clrs[4], alpha = 0.75),
cells_body(columns = value_linear, 
rows = (upper_linear < 0) | (lower_linear > 0))) |>
tab_style(cell_fill(color = clrs[4], alpha = 0.75),
cells_body(columns = value_quadratic, 
rows = (upper_quadratic < 0) | (lower_quadratic > 0))) |>
tab_style(cell_fill(color = clrs[4], alpha = 0.75),
cells_body(columns = value_cubic, 
rows = (upper_cubic < 0) | (lower_cubic > 0))) |> 
tab_style(cell_text(size = "small"), 
list(cells_body(), cells_stub())) |> 
tab_footnote("Sum-coded contrast between 21 and 25 month old participants.",
cells_stub(rows = 2)) |> 
tab_footnote("Sum-coded contrast between 25 and 30 month old participants.",
cells_stub(rows = 3)) |> 
tab_footnote("Sum-coded contrast between monolingual and bilingual participants.",
cells_stub(rows = 4)) |> 
tab_footnote("Sum-coded contrast between related (cognate and non-cognate) and unrelated trials",
cells_stub(rows = 5)) |> 
tab_footnote("Sum-coded contrast between cognate and non-cognate trials",
cells_stub(rows = 6))
```

#### Plot

```{r}
#| label: fig-fixed-effects
#| fig-height: 7
#| fig-width: 13
#| fig-cap: "Summary of the posterior of the Growth Curve Analysis model. We report the median and 95% highest density interval (*HDI*) of each fixed coefficient. The 95% interval indicates the range of values that contains the true value of the parameter with 95% probability. Values are in *logit* scale. Coefficients whose 95% HDI excludes zero are highlighted in red. Summary of the posterior distribution of the Growth Curve Analysis model. We report the median and 95% highest density interval (*HDI*) of each fixed coefficient. The 95% interval indicates the range of values that contains the true value of the parameter with 95% probability. Values are in *logit* scale. Coefficients whose 95% HDI excludes zero are highlighted in red. Coefficients on the intercept indicate changes in overal proportion of target looking, coefficients in the linear term indicate changes in the overall slope of target looking growth across time, coefficients in the quadratic and cubic terms reflect the steepness of the rise and fall or of the rise, fall and rise, respectivey,  of the curve."
post_draws |> 
median_hdi() |>
mutate(poly = case_when(str_detect(.variable, "ot1") ~ "Linear",
str_detect(.variable, "ot2") ~ "Quadratic",
str_detect(.variable, "ot3") ~ "Cubic",
TRUE ~ "Intercept") |> 
factor(levels = c("Intercept", "Linear", "Quadratic", "Cubic")),
.variable_name = case_when(.variable %in% paste0("b_ot", 1:3) ~ .variable,
TRUE ~ str_remove_all(.variable, "ot1\\:|ot2\\:|ot3\\:")) |> 
factor(levels = names(str_repl),
label = str_repl,
ordered = TRUE),
type = case_when(.variable_name %in% str_repl[1:9] ~ "Main effect",
.variable_name %in% str_repl[10:17] ~ "Two-way interactions",
.variable_name %in% str_repl[18:length(str_repl)] ~ "Three-way interactions"),
excludes_zero = ifelse((.lower > 0) | (.upper < 0), "Yes", "No")) |>
select(.variable_name, type, poly, .value:.upper, excludes_zero) |>
arrange(.variable_name) |> 
ggplot(aes(x = .value, y = fct_rev(.variable_name), 
colour = excludes_zero,
xmin = .lower, xmax = .upper)) +
facet_wrap(~poly, scales = "free_x", nrow = 1) +
geom_vline(xintercept = 0, colour = "grey", size = 1) +
geom_errorbar(width = 0.2, size = 0.75) +
geom_point(size = 3) +
labs(x = "Estimate (Logit scale)", 
y = "Variable", 
fill = "CrI", 
alpha = "CrI",
colour = "Interval excludes zero?") +
theme(legend.position = "top",
axis.title.y = element_blank(),
axis.text.y = element_text(face = "plain", hjust = 1),
panel.border = element_rect(colour = "black", fill = NA))
```

@@

```{r}
#| label: fig-posterior-epred
#| fig-width: 12
#| fig-height: 8
nd <- expand_grid(timebin = seq(0, 20, 0.1),
age_group = levels(data_time$age_group),
lp = levels(data_time$lp),
trial_type = levels(data_time$trial_type)
) |> 
polypoly::poly_add_columns(timebin,
degree = 3,
prefix = "ot",
scale_width = 1)

@@

m <- add_epred_draws(nd,
fit_timecourse_2 ,
ndraws = 200,
re_formula = NA) |> 
mutate(.epred = plogis(.epred))

ptlt_means <- data_time |> 
summarise(.prop = mean(.prop), 
.by = c(id, age_group, lp, trial_type, timebin)) |> 
summarise(.prop = mean(.prop),
.by = c(age_group, lp, trial_type, timebin))

m |> 
ggplot(aes(timebin,
Estimate,
colour = trial_type,
fill = trial_type)) +
facet_grid(lp~age_group) +
geom_hline(yintercept = 0.5, 
size = 1,
colour = "grey") +
geom_ribbon(aes(ymin = Q2.5,
ymax = Q97.5),
colour = NA,
alpha = 1/3) +
geom_line(linewidth = 3/4) +
# stat_summary(fun = median, geom = "line", size = 1) +
geom_point(data = ptlt_means, aes(y = .prop)) +
labs(x = "Time (ms)",
y = "Proportion of target looking time",
colour = "Prime",
fill = "Prime",
shape = "Prime",
linetype = "Trial type") +
scale_x_continuous(labels = \(x) format(250 + x*100, big.mark = ",")) +
scale_y_continuous(limits = c(0, 1)) +
theme(legend.position = "top",
panel.grid = element_line(linetype = "dotted", colour = "grey"),
panel.grid.minor = element_blank(),
legend.title = element_blank())
```

@@

>>>>

# load target objects
tar_load(c(participants, stimuli, 
		   attrition_trials,
		   attrition_participants,
		   gaze_aoi, vocabulary,
		   loos_aggregated, loos_timecourse,
		   data_time, data_summary))

@@

```{r attrition-numbers}
n_trials <- nrow(attrition_trials)
n_sessions <- distinct(attrition_trials, id, age_group) |> 
	nrow()
n_participants <- distinct(attrition_trials, id) |> 
	nrow()

n_trials_prime <- attrition_trials |> 
	filter(is_valid_gaze_prime)

n_trials_test <- n_trials_prime |> 
	filter(is_valid_gaze_test)

n_trials_each <- n_trials_test |> 
	filter(is_valid_gaze_test_each)

n_trials_participant <- n_trials_each |> 
	inner_join(filter(attrition_participants,
					  is_valid_participant),
			   by = join_by(id, age_group))

n_sessions_valid <- attrition_participants |> 
	filter(is_valid_participant)

n_participants_valid <- n_trials_participant |> 
	distinct(id)

n_longitudinal <- count(n_sessions_valid, id, name = "sessions") |> 
	count(sessions)

n_participants_include <- read_csv("data-raw/participants.csv",
								   show_col_types = FALSE)
```

We gathered data for `r format(n_trials, big.mark = ",")` trials from `r n_sessions` experimental sessions, provided by `r n_participants` distinct participants. Missing eye-tracker samples were interpolated using the last-observation-carried-forward [see @zettersten2022peekbank for a similar approach]. We excluded trials in which the participant did not fixate the prime at least half of the time of the prime phase ($\geq$750 ms, *n* = `r format(n_trials - nrow(n_trials_prime), big.mark = ",")`), or the target and distractor during the test phase ($\geq$1,000 ms, *n* = `r format(nrow(n_trials_prime) - nrow(n_trials_test), big.mark = ",")`). Additionally, we also excluded trials in which the participant failed to fixate both target *and* distractor pictures for at least 100 ms (5% of the time, `r format(nrow(n_trials_test) - nrow(n_trials_each), big.mark = ",")`). Finally we excluded trials from experimental sessions in which the participant did not provide at least two valid trials in each of the three experimental conditions (cognate prime, non-cognate prime, and unrelated prime) which resulted in the exclusion of `r nrow(distinct(n_trials_each, id, age_group)) - nrow(n_sessions_valid)` experimental sessions and their `r nrow(n_trials_each) - nrow(n_trials_participant)` remaining trials. The final data set comprised `r format(sum(data_summary$.ntrials), big.mark = ",")` from `r n_sessions_valid` experimental sessions, provided by `r n_participants_valid` distinct participants. Of those participants, `r n_longitudinal[1]` provided data from one experimental session, `r n_longitudinal[2]` provided data from two experimental sessions, and `r n_longitudinal[1]` provided data from three experimental sessions.

We defined our time window of interest from 300 ms after the onset of the test phase (target and distractor presentation) until the end of the test phase (2,000 ms). For each trial, we chunked the time domain into 17 time bins of 100 ms of duration. We then calculated, for each experimental session, time bin, and condition, participant's proportion of target and distractor fixations. Finally, we computed the empirical log-odds of target and distractor fixations, which we introduced in the statistical analyses as our response variable.

@@

@@

## Aggregated target looking

```{r}
#| label: tbl-loos-agreggated
loos_aggregated |> 
	as_tibble(.name_repair = janitor::make_clean_names) |> 
	rownames_to_column("model") |> 
	relocate(
		model, 
		matches("diff"),
		matches("elpd_loo"),
		matches("p_loo"), 
		matches("looic")
	) |>
	mutate(across(everything(), as.numeric),
		   model = attr(loos_aggregated, "dimnames")[[1]]) |> 
	select(-matches("subsampling")) |> 
	gt(rowname_col = "model") |>
	fmt_number(is.numeric, decimals = 1) |> 
	cols_merge_uncert(col_val = elpd_diff, col_uncert = se_diff) |> 
	cols_merge_uncert(col_val = elpd_loo, col_uncert = se_elpd_loo) |> 
	cols_merge_uncert(col_val = p_loo, col_uncert = se_p_loo) |> 
	cols_merge_uncert(col_val = looic, col_uncert = se_looic) |> 
	cols_label(model = "Model",
			   elpd_loo = md("*ELPD*"),
			   p_loo = md("*p*"),
			   looic = md("*LOO-IC*"),
			   elpd_diff = md("Difference")) |> 
	tab_source_note(md("Pareto-*k* estimates of all models were acceptable (*k* < 0.5)"))

## Growth Curve Analysis

```{r}
#| label: tbl-loos-timecourse
loos_timecourse |> 
	as_tibble(.name_repair = janitor::make_clean_names) |> 
	rownames_to_column("model") |> 
	relocate(
		model, 
		matches("diff"),
		matches("elpd_loo"),
		matches("p_loo"), 
		matches("looic")
	) |>
	mutate(across(everything(), as.numeric),
		   model = attr(loos_timecourse, "dimnames")[[1]]) |> 
	select(-matches("subsampling")) |> 
	gt(rowname_col = "model") |>
	fmt_number(is.numeric, decimals = 1) |> 
	cols_merge_uncert(col_val = elpd_diff, col_uncert = se_diff) |> 
	cols_merge_uncert(col_val = elpd_loo, col_uncert = se_elpd_loo) |> 
	cols_merge_uncert(col_val = p_loo, col_uncert = se_p_loo) |> 
	cols_merge_uncert(col_val = looic, col_uncert = se_looic) |> 
	cols_label(model = "Model",
			   elpd_loo = md("*ELPD*"),
			   p_loo = md("*p*"),
			   looic = md("*LOO-IC*"),
			   elpd_diff = md("Difference")) |> 
	tab_source_note(md("Pareto-*k* estimates of all models were acceptable (*k* < 0.5)"))

@@

```{r}
#| label: fig-posterior-epred
#| fig-width: 12
#| fig-height: 8
nd <- expand_grid(timebin = seq(1, 17, 0.1),
				  age_group = levels(data_time$age_group),
				  lp = levels(data_time$lp),
				  trial_type = levels(data_time$trial_type)) |> 
	polypoly::poly_add_columns(timebin,
							   degree = 3,
							   prefix = "ot",
							   scale_width = 1)

@@

# m <- add_epred_draws(nd,
# 					 fit_timecourse_2 ,
# 					 ndraws = 200,
# 					 re_formula = NA) |> 
# 	mutate(.epred = plogis(.epred))

ptlt_means <- data_time |> 
	summarise(.logit = mean(.logit), 
			  .by = c(id, age_group, lp, trial_type, timebin)) |> 
	summarise(.logit = mean(.logit),
			  .by = c(age_group, lp, trial_type, timebin))

m |> 
	ggplot(aes(timebin,
			   Estimate,
			   colour = trial_type,
			   fill = trial_type)) +
	facet_grid(lp~age_group) +
	geom_hline(yintercept = 0.5, 
			   size = 1,
			   colour = "grey") +
	geom_ribbon(aes(ymin = Q2.5,
					ymax = Q97.5),
				colour = NA,
				alpha = 1/3) +
	geom_line(linewidth = 3/4) +
	stat_summary(fun = median, geom = "line", size = 1) +
	geom_point(data = ptlt_means,
			   aes(y = plogis(.logit))) +
	labs(x = "Time (ms)",
		 y = "Proportion of target looking time",
		 colour = "Prime",
		 fill = "Prime",
		 shape = "Prime",
		 linetype = "Trial type") +
	scale_x_continuous(labels = \(x) format(250 + x*100, big.mark = ",")) +
	scale_y_continuous(limits = c(0, 1)) +
	theme(legend.position = "top",
		  panel.grid = element_line(linetype = "dotted", colour = "grey"),
		  panel.grid.minor = element_blank(),
		  legend.title = element_blank())
```

@@

## Trial-level attrition by participant

### Trial-level attrition by stimuli properties
